---
title: "Extended Language Network | 02b_stats_selectivityFallOff"
author: "Agata Wolna [`awolna@mit.edu`], Aaron Wright [`aawva@mit.edu`]"
date: "2026-06-24"
format: html
---

## Overview

**Purpose**: Run statistical models evaluating responsiveness and selectivity for language across Harvard-Oxford subcortical parcels using the selectivity fall-off dataset (fROI defined at top 10% threshold).

**Input(s)**: `data/detailed_selectivity_falloff_all.txt` — per-subject effect sizes across fROI thresholds and conditions

**Process**: Fit mixed-effects models (lmer) per subcortical ROI testing differences between conditions and tasks (Lang, Alice, MD localizers); run emmeans post-hoc contrasts to identify language-responsive and language-selective parcels.

**Output(s)**: `results/stats/selectivityFallOff_model_stats/`:
- `subcort_model_all_effects.txt` — main model fixed effects per ROI
- `subcort_model_pairwise_condition.txt` — critical vs. baseline contrasts per task
- `subcort_model_pairwise_task.txt` — between-task contrasts at critical conditions
- `subcort_model_pairwise_maineffects.txt` — condition-vs-zero contrasts
- `subcort_model_all_means.txt` — estimated marginal means per task × condition
- `subcort_responsive_rois.txt` — ROIs classified as language-responsive
- `subcort_selective_rois.txt` — ROIs classified as language-selective

## Environment

```{r}
library(tidyverse)
library(here)
library(lmerTest)
library(emmeans)

source(here("scripts", "run_parcel_models.R"))
```

## Data

```{r}
data_selectivity_falloff <- read.delim(here("data", "detailed_selectivity_falloff_all.txt")) %>%
  as_tibble() %>%
  filter(included == 1)

data_subcort <- data_selectivity_falloff %>%
  filter(parcels == "HOSubcortical", !is.na(condition))

# n check
data_subcort %>%
  select(Subject, task, localizer) %>%
  unique() %>%
  count(task, localizer)
```

## Tidying

```{r}
# classify conditions as critical or baseline
baseline <- c("easy vWM", "nonwords", "degraded speech")

dat <- data_subcort %>%
  mutate(
    cond = ifelse(condition %in% baseline, "baseline", "critical"),
    cond = as.factor(cond),
    task = as.factor(task))

# set contrast coding: treatment contrasts
contrasts(dat$cond) <- contr.treatment(2)
contrasts(dat$task) <- contr.treatment(3, base = 1)

rois  <- dat$ROI %>% unique()
tasks <- c("Alice", "Lang", "MD")
```

## Statistical models: subcortical parcels

Models are fit per ROI on data from the Lang localizer at the top 10% fROI threshold, testing differences between conditions and tasks.

```{r}
# initialise output tables
roi_models_subcort_all             <- tibble(matrix(ncol = 7, nrow = 0))
roi_models_subcort_pairwise_condition  <- tibble(matrix(ncol = 8, nrow = 0))
roi_models_subcort_pairwise_task       <- tibble(matrix(ncol = 8, nrow = 0))
roi_models_subcort_pairwise_maineffects <- tibble(matrix(ncol = 8, nrow = 0))
roi_models_subcort_all_means       <- tibble(matrix(ncol = 8, nrow = 0))

for (i in rois) {
  # model: condition × task
  model_subcort <- lmer(
    EffectSize ~ 1 + cond + task + cond:task + (1 | Subject),
    data    = dat %>% filter(thr == "top10", ROI == i, localizer == "Lang"),
    control = lmerControl(optimizer = "bobyqa"))

  # fixed effects summary
  sum_coef <- summary(model_subcort)$coef
  sum_coef <- cbind(rownames(sum_coef), sum_coef, i) %>% as_tibble()
  roi_models_subcort_all <- roi_models_subcort_all %>% rbind(sum_coef)

  # estimated marginal means
  emm <- emmeans(model_subcort, ~ task * cond)

  # condition vs. zero (each cell vs. 0)
  cond_vs_zero <- contrast(emm, method = list(
    "Alice baseline"    = c(1, 0, 0, 0, 0, 0),
    "Alice critical"    = c(0, 0, 0, 1, 0, 0),
    "Language baseline" = c(0, 1, 0, 0, 0, 0),
    "Language critical" = c(0, 0, 0, 0, 1, 0),
    "MD baseline"       = c(0, 0, 1, 0, 0, 0),
    "MD critical"       = c(0, 0, 0, 0, 0, 1)),
    adjust = "none") %>%
    as.data.frame() %>% as_tibble() %>% cbind(i)

  # between-task contrasts at critical conditions
  critical_between_tasks <- contrast(emm, method = list(
    "criticalAlice vs. criticalLang" = c(0, 0, 0,  1, -1,  0),
    "criticalAlice vs. criticalMD"   = c(0, 0, 0,  1,  0, -1),
    "criticalLang vs. criticalMD"    = c(0, 0, 0,  0,  1, -1)),
    adjust = "none") %>%
    as.data.frame() %>% as_tibble() %>% cbind(i)

  # critical vs. baseline per task
  critical_vs_baseline <- contrast(emm, method = list(
    "Alice critical vs. baseline" = c(-1, 0, 0, 1, 0, 0),
    "Lang critical vs. baseline"  = c(0, -1, 0, 0, 1, 0),
    "MD critical vs. baseline"    = c(0,  0, -1, 0, 0, 1)),
    adjust = "none") %>%
    as.data.frame() %>% as_tibble() %>% cbind(i)

  means <- emm %>% as.data.frame() %>% cbind(i)

  roi_models_subcort_pairwise_condition   <- roi_models_subcort_pairwise_condition   %>% rbind(critical_vs_baseline)
  roi_models_subcort_pairwise_task        <- roi_models_subcort_pairwise_task        %>% rbind(critical_between_tasks)
  roi_models_subcort_pairwise_maineffects <- roi_models_subcort_pairwise_maineffects %>% rbind(cond_vs_zero)
  roi_models_subcort_all_means            <- roi_models_subcort_all_means            %>% rbind(means)
}

# rename columns
colnames(roi_models_subcort_all)              <- c("effect", "estimate", "se", "df", "t_value", "p_value", "roi")
colnames(roi_models_subcort_pairwise_condition)   <- c("contrast", "estimate", "se", "df", "z_ratio", "p_value", "roi")
colnames(roi_models_subcort_pairwise_task)        <- c("contrast", "estimate", "se", "df", "z_ratio", "p_value", "roi")
colnames(roi_models_subcort_pairwise_maineffects) <- c("contrast", "estimate", "se", "df", "z_ratio", "p_value", "roi")
colnames(roi_models_subcort_all_means)        <- c("task", "condition", "emmean", "se", "df", "lowCI", "highCI", "roi")

# attach ROI labels
labels <- dat %>% select(ROI, label) %>% unique() %>% rename(roi = ROI)

roi_models_subcort_pairwise_condition   <- roi_models_subcort_pairwise_condition   %>% left_join(labels)
roi_models_subcort_pairwise_task        <- roi_models_subcort_pairwise_task        %>% left_join(labels)
roi_models_subcort_pairwise_maineffects <- roi_models_subcort_pairwise_maineffects %>% left_join(labels)
roi_models_subcort_all_means            <- roi_models_subcort_all_means            %>% left_join(labels)
```

## Responsive and selective parcels

```{r}
# language-responsive: significant critical > baseline for Alice or Lang,
# AND significant critical effect vs. zero
responsive_contrast <- roi_models_subcort_pairwise_condition %>%
  filter(contrast != "MD critical vs. baseline", p_value < .05) %>%
  select(roi, label) %>%
  unique() %>%
  mutate(responsive_contrast = 1)

responsive_non_zero <- roi_models_subcort_pairwise_maineffects %>%
  filter(contrast %in% c("Alice critical", "Language critical"), estimate > 0, p_value < .05) %>%
  select(roi, label) %>%
  unique() %>%
  mutate(responsive_non_zero = 1)

responsive <- responsive_contrast %>%
  left_join(responsive_non_zero, by = c("roi", "label")) %>%
  filter(responsive_contrast == responsive_non_zero)

# language-selective: significantly greater critical response for Lang/Alice than MD
selective_candidates <- roi_models_subcort_pairwise_task %>%
  filter(contrast != "criticalAlice vs. criticalLang", p_value < .05, estimate > 0) %>%
  select(roi, label) %>%
  unique() %>%
  mutate(selective = 1)

selective <- responsive %>%
  left_join(selective_candidates, by = c("roi", "label")) %>%
  filter(!is.na(selective)) %>%
  select(-selective)
```

## Save results

```{r}
dir.create(here("results", "stats", "selectivityFallOff_model_stats"), recursive = TRUE, showWarnings = FALSE)

write.table(roi_models_subcort_all,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_model_all_effects.txt"),
  sep = "\t", row.names = FALSE)

write.table(roi_models_subcort_pairwise_condition,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_model_pairwise_condition.txt"),
  sep = "\t", row.names = FALSE)

write.table(roi_models_subcort_pairwise_task,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_model_pairwise_task.txt"),
  sep = "\t", row.names = FALSE)

write.table(roi_models_subcort_pairwise_maineffects,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_model_pairwise_maineffects.txt"),
  sep = "\t", row.names = FALSE)

write.table(roi_models_subcort_all_means,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_model_all_means.txt"),
  sep = "\t", row.names = FALSE)

write.table(responsive,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_responsive_rois.txt"),
  sep = "\t", row.names = FALSE)

write.table(selective,
  here("results", "stats", "selectivityFallOff_model_stats", "subcort_selective_rois.txt"),
  sep = "\t", row.names = FALSE)
```
