---
title: "Extended Language Network | 03_plots"
author: "Aaron Wright [`aawva@mit.edu`], Agata Wolna [`awolna@mit.edu`]"
date: "2025-08-29"
format: html
---

## Overview

**Purpose**: create various plots depicting the distribution of language-responsive and language-selective cortical and subcortical ROIs

**Input(s)**: `AtlasParcels.xlsx`; `atlas_roi_models_pairwise_conditions.txt`, `gss_n772_roi_models_all_means.txt`, `gss_n86_roi_models_all_means.txt` and the lists of lang-responsive and lang-selective parcels (all produced in 00_stats); % overlap values between the atlas and functional language parcels (`voxel_overlap_results_average.txt`)

**Process**: import and wrangle model statistical data and create bar descriptive bar plots and brain atlas plots using `{ggseg}`

**Output(s)**: parcel activation bar plots; by-roi inset figures for the n=772 and n=86 gss models, and atlas parcels models; the base for the main cortical and subcortical figures from the paper; and bar plots showing the overlap between atlas and functional language parcels

## Environment

```{r}
library(tidyverse)
library(here)
library(readxl)
library(ggsflabel)
library(ggseg)
library(ggsegGlasser)
library(ggsegDKT)
library(ggsegHO)
library(patchwork)

source(here("scripts", "plotting_aesthetics.R")) # EvLab contrast aesthetics
source(here("scripts", "plot_brains.R")) # functions for batch plotting atlas figures
```

## Data

### Raw data

```{r}
data_all_recoded_gss_full <-  read_csv(here("results", "tables", "gss_raw_data_detail_all_recoded_descriptive_stats.csv")) # ALL participants' raw data, with parcels attached
```

### Statistical models

```{r}
# read in data

# statistical model data 
# GSS model - full model
gss_means_plots_n772 <- read.delim(here("results", "stats", "gss_n772_model_stats", "gss_n772_roi_models_all_means.txt")) %>% as_tibble() %>% select(-df)

# Atlas model - main model
roi_models_atlas <- read.delim(here("results", "stats", "atlas_model_stats", "atlas_roi_models_pairwise_conditions.txt"), sep = "\t") %>% as_tibble()

# ROIs language-responsive and language-selective
lang_responsive_rois <- read.delim(here("results", "tables", "atlas_lang_responsive_rois_names.txt"), sep = "\t") %>% as_tibble() %>% mutate(lang_resp = 1) %>% select(-hemi, -ROIName)
lang_selective_rois <- read.delim(here("results", "tables", "atlas_lang_selective_rois_names.txt"), sep = "\t") %>% as_tibble() %>% mutate(lang_sel = 1) %>% select(-hemi, -ROIName)
```

## Tidying (statistical model data: GSS and atlases)

```{r}
# create a column that assigns 1 to language selective and responsive parcels and 0 to the remaining parcels
roi_models_atlas <- roi_models_atlas %>%
  left_join(lang_responsive_rois) %>%
  left_join(lang_selective_rois) %>%
  separate(contrast, into = "task") %>%
  rename(parcels = atlas)

# read in subcortical data
hosubcortical_parcels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet ="HOSubcortical") %>% rename(roi=ROI)

# Atlas parcels (cortical) overlap with GcSS parcels
atlas_overlap_average <- read.delim(here("data", "voxel_overlap_results_average.txt")) %>%
  as_tibble() %>%
  rename(roi = ROI_Atlas) %>% 
  mutate(parcels = ifelse(Atlas == "HOCortical_2mm.nii", "HOC", ifelse(Atlas == "DKT_in_MNI_sm2.nii", "DKT", "Glasser"))) %>% 
  select(-Atlas, -ROI_GcSS)

# create a data frame to define task effects
effects <- data.frame(
  column1 = c("Alice", "Lang", "MD", "Alice", "Lang", "MD"),
  column2 = c("baseline", "baseline", "baseline", "critical", "critical", "critical"),
  column3 = c("degraded speech", "nonwords", "easy vWM", "intact speech", "sentences", "hard vWM"))
# rename columns
colnames(effects) <- c("task", "condition", "effect")

# read in parcel decoding files
parcels <- read.delim(here("data", "recode_parcels.txt"), sep = "\t") %>% select(-ROI) %>% rename("roi" = "ROI_new")

# join effects + gss_means_plots - n772
gss_means_plots_n772 <- gss_means_plots_n772 %>%
  left_join(effects, by = c("task", "condition")) %>%
  left_join(parcels, by = "roi") %>%
  mutate(model = "n=772")
```

## Pre-processing atlas data

### Desikan-Killiany-Tourville (DKT) atlas

```{r}
# get DKT labels
dktlabels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet = "DKT") %>% rename(roi = ROI, region = ROIName)
dktparc <- dkt %>% as.data.frame() %>% select(hemi, region, side) %>% distinct()

## preparing data
dkt_plot_lang <- roi_models_atlas %>% filter(parcels == "DKT") %>% left_join(dktlabels) %>% rename(MeanEffect = estimate)

dkt_plot_lang <- dktparc %>% left_join(dkt_plot_lang, relationship = "many-to-many") %>% as_tibble() %>% 
  mutate(lang_resp = ifelse(is.na(lang_resp), 0, 1), 
         lang_sel = ifelse(is.na(lang_sel),0, 1), 
         MeanEffect = ifelse(is.na(MeanEffect), 0, MeanEffect),
         MeanEffectSelect = MeanEffect*lang_sel,
         MeanEffectSelect = ifelse(MeanEffectSelect<0, 0, MeanEffectSelect),
         MeanEffectResponsive = MeanEffect*lang_resp,
         MeanEffectResponsive = ifelse(MeanEffectResponsive<0, 0, MeanEffectResponsive),
         sig = as.factor(lang_sel)) %>%
  ### dkt atlas contains 'roi' column of type 'character' - no other atlas inherently includes this column, and removing solves conflict when plotting overlap
  select(-roi) %>% 
  filter(!is.na(region))

## add columns with % overlap with functional (GcSS) parcels
dkt_dat_overlap <- atlas_overlap_average %>% filter(parcels=="DKT") %>% left_join(dktlabels)
# select -roi for consistency
dkt_dat_overlap <- dktparc %>% drop_na(region) %>% left_join(dkt_dat_overlap) %>% as_tibble() %>% select(-c(parcels, roi))

dkt_plot_lang <- dkt_dat_overlap %>% left_join(dkt_plot_lang) 

## count language-selective and language-responsive parcels
dkt_plot_lang %>% filter(lang_sel == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)
dkt_plot_lang %>% filter(lang_resp == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)

# save a dataframe with DKT parcels names and info on language responsiveness and selectivity
dkt_roi_classification <- dkt_plot_lang %>% select(parcels, hemi, region, side, lang_resp, lang_sel)
outfile <- here("results", "tables", "dkt", "dkt_lang_responsive_selective_rois.txt")
dir.create(dirname(outfile), recursive = TRUE, showWarnings = FALSE)

# save dataframe
write.table(dkt_roi_classification, outfile, sep = "\t", row.names = FALSE, quote = FALSE)
```

### Harvard-Oxford Cortical atlas

```{r}
# fix the Angular Gyrus and Supramarginal Gyrus post. geometries - error (reversed order) is internal to the hoCort ggseg data
## retrieve geometries
lh_angular_geometry <- hoCort$data$geometry[1]
lh_supramarginal_post_geometry <- hoCort$data$geometry[31]
## reassign
hoCort$data$geometry[1] <- lh_supramarginal_post_geometry
hoCort$data$geometry[31] <- lh_angular_geometry

# get HOC (Harvard-Oxford Cortical) labels
hoclabels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet = "HOCortical") %>% rename(roi = ROI, region = ROIName)
hocparc <- hoCort %>% as.data.frame() %>% select(hemi, region, side) %>% distinct()

## fix the Heschl gyrus labels
hoclabels$region[45] <- hocparc$region[6]
hoclabels$region[93] <- hocparc$region[6]

## preparing data
hoCort_plot_lang <- roi_models_atlas %>% filter(parcels == "HOCortical") %>% left_join(hoclabels) %>% rename(MeanEffect = estimate)
hoCort_plot_lang <- hocparc %>% left_join(hoCort_plot_lang, relationship = "many-to-many") %>% as_tibble() %>%
  mutate(lang_resp = ifelse(is.na(lang_resp), 0, 1), 
         lang_sel = ifelse(is.na(lang_sel),0, 1), 
         MeanEffect = ifelse(is.na(MeanEffect), 0, MeanEffect),
         MeanEffectSelect = MeanEffect*lang_sel,
         MeanEffectSelect = ifelse(MeanEffectSelect<0, 0, MeanEffectSelect),
         MeanEffectResponsive = MeanEffect*lang_resp,
         MeanEffectResponsive = ifelse(MeanEffectResponsive<0, 0, MeanEffectResponsive),
         sig = as.factor(lang_sel)) %>% 
  filter(!is.na(region))

## add columns with % overlap with functional (GcSS) parcels
hoCort_dat_overlap <- atlas_overlap_average %>% filter(parcels == "HOC") %>% left_join(hoclabels)
hoCort_dat_overlap <- hocparc %>% drop_na(region) %>% left_join(hoCort_dat_overlap) %>% as_tibble() %>% select(-parcels)

hoCort_plot_lang <- hoCort_dat_overlap %>% left_join(hoCort_plot_lang)

## count language-selective and language-responsive parcels
hoCort_plot_lang %>% filter(lang_sel == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)
hoCort_plot_lang %>% filter(lang_resp == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)

# save a dataframe with HOC parcels names and info on language responsiveness and selectivity
hoCort_roi_classification <- hoCort_plot_lang %>% select(parcels, hemi, region, side, lang_resp, lang_sel)
outfile <- here("results", "tables", "hoCort", "hoCort_lang_responsive_selective_rois.txt")
dir.create(dirname(outfile), recursive = TRUE, showWarnings = FALSE)

# save dataframe
write.table(hoCort_roi_classification, outfile, sep = "\t", row.names = FALSE, quote = FALSE)

```

### Glasser atlas

```{r}
# get Glasser labels
glasserlabels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet = "Glasser") %>% rename(roi = ROI, region = ROIName)
glasserparc <- glasser %>% as.data.frame() %>% select(hemi, region, side) %>% distinct()

## preparing data
glasser_plot_lang <- roi_models_atlas %>% filter(parcels == "Glasser") %>% left_join(glasserlabels) %>% rename(MeanEffect = estimate)

glasser_plot_lang <- glasserparc %>% left_join(glasser_plot_lang, relationship = "many-to-many") %>% as_tibble() %>% 
  mutate(lang_resp = ifelse(is.na(lang_resp), 0, 1), 
         lang_sel = ifelse(is.na(lang_sel),0, 1), 
         MeanEffect = ifelse(is.na(MeanEffect), 0, MeanEffect),
         MeanEffectSelect = MeanEffect*lang_sel,
         MeanEffectSelect = ifelse(MeanEffectSelect<0, 0, MeanEffectSelect),
         MeanEffectResponsive = MeanEffect*lang_resp,
         MeanEffectResponsive = ifelse(MeanEffectResponsive<0, 0, MeanEffectResponsive),
         sig = as.factor(lang_sel)) %>%
  select(-roi) %>% 
  filter(!is.na(region))

## add columns with % overlap with functional (GcSS) parcels
glasser_dat_overlap <- atlas_overlap_average %>% filter(parcels == "Glasser") %>% left_join(glasserlabels)
glasser_dat_overlap <- glasserparc %>% drop_na(region) %>% left_join(glasser_dat_overlap) %>% as_tibble() %>% select(-parcels)

glasser_plot_lang <- glasser_dat_overlap %>% left_join(glasser_plot_lang)

## count language-selective and language-responsive parcels
glasser_plot_lang %>% filter(lang_sel == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)
glasser_plot_lang %>% filter(lang_resp == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)

# save a dataframe with Glasser parcels names and info on language responsiveness and selectivity
glasser_roi_classification <- glasser_plot_lang %>% select(parcels, hemi, region, side, lang_resp, lang_sel)
outfile <- here("results", "tables", "glasser", "glasser_lang_responsive_selective_rois.txt")
dir.create(dirname(outfile), recursive = TRUE, showWarnings = FALSE)

# save dataframe
write.table(glasser_roi_classification, outfile, sep = "\t", row.names = FALSE, quote = FALSE)
```

### Harvard-Oxford Subcortical atlas

As of November 2025, there is no published ggseg atlas for the Harvard-Oxford Subcortical atlas. We made use of the Aseg subcortical atlas from the {ggsegExtra} package, found [here](https://ggseg.github.io/ggsegExtra/articles/createaseg.html).

```{r}
# get HOSC (Harvard-Oxford Subcortical) labels
hosclabels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet="HOSubcortical") %>% mutate(lobe = NA) %>% rename(roi = ROI, region = ROIName)
## preprocessing aseg parcels to align with hosc labels for plotting
hoscparc <- aseg %>% 
  as.data.frame() %>% 
  select(hemi, region, side) %>% 
  distinct() %>%
  mutate(original_region = region,  # Store original region names temporarily
         region = case_when(
           region == "thalamus proper" & hemi == "left" ~ "Left Thalamus",
           region == "thalamus proper" & hemi == "right" ~ "Right Thalamus",
           region == "lateral ventricle" & hemi == "left" ~ "Left Lateral Ventricle",
           region == "lateral ventricle" & hemi == "right" ~ "Right Lateral Ventricle",
           region == "caudate" & hemi == "left" ~ "Left Caudate",
           region == "caudate" & hemi == "right" ~ "Right Caudate",
           region == "putamen" & hemi == "left" ~ "Left Putamen",
           region == "putamen" & hemi == "right" ~ "Right Putamen",
           region == "hippocampus" & hemi == "left" ~ "Left Hippocampus",
           region == "hippocampus" & hemi == "right" ~ "Right Hippocampus",
           region == "pallidum" & hemi == "left" ~ "Left Pallidum",
           region == "pallidum" & hemi == "right" ~ "Right Pallidum",
           region == "amygdala" & hemi == "left" ~ "Left Amygdala",
           region == "amygdala" & hemi == "right" ~ "Right Amygdala",
           TRUE ~ region  # Keep all other regions unchanged
         )) %>%
  left_join(hosclabels, by = "region") %>%
  mutate(hemi = coalesce(hemi.x, hemi.y)) %>%
  select(hemi, original_region, region, side, roi)

## preparing data
hoSubCort_plot_lang <- roi_models_atlas %>% filter(parcels == "HOSubcortical") %>% left_join(hosclabels) %>% rename(MeanEffect = estimate)
hoSubCort_plot_lang <- hoscparc %>% left_join(hoSubCort_plot_lang) %>% as_tibble() %>%
  # reverse the renaming after merge to get back to the original region names
  select(-region) %>%
  rename(region = original_region) %>%
  mutate(lang_resp = ifelse(is.na(lang_resp), 0, 1), 
         lang_sel = ifelse(is.na(lang_sel),0, 1), 
         MeanEffect = ifelse(is.na(MeanEffect), 0, MeanEffect),
         MeanEffectSelect = MeanEffect*lang_sel,
         MeanEffectSelect = ifelse(MeanEffectSelect<0, 0, MeanEffectSelect),
         MeanEffectResponsive = MeanEffect*lang_resp,
         MeanEffectResponsive = ifelse(MeanEffectResponsive<0, 0, MeanEffectResponsive),
         sig = as.factor(lang_sel)) %>% 
  filter(!is.na(region))

## count language-selective and language-responsive parcels
hoSubCort_plot_lang %>% filter(lang_sel == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)
hoSubCort_plot_lang %>% filter(lang_resp == 1) %>% select(hemi, region) %>% unique() %>% count(hemi)

# save a dataframe with HOSC parcels names and info on language responsiveness and selectivity
hoSubCort_roi_classification <- hoSubCort_plot_lang %>% select(parcels, hemi, region, side, lang_resp, lang_sel)
outfile <- here("results", "tables", "hoSubCort", "hoSubCort_lang_responsive_selective_rois.txt")
dir.create(dirname(outfile), recursive = TRUE, showWarnings = FALSE)

# save dataframe
write.table(hoSubCort_roi_classification, outfile, sep = "\t", row.names = FALSE, quote = FALSE)
```

## Plotting: raw data by-ROI bar plots

### GSS cortical by-roi bar plots (raw data)

```{r}
# read in the data
data_all_recoded_gss_full <- read_csv(here("results", "tables", "gss_raw_data_detail_all_recoded_descriptive_stats.csv")) # ALL participants' raw data, with parcels attached

# define config for raw pipeline
raw_plot_config <- list(
  df = data_all_recoded_gss_full,
  folder = "gss_n772_raw_data_rois",
  combined_file = "GSS_n772_raw_language_alice_md_left_cortical_combined.jpg"
)

df <- raw_plot_config$df

# roi labels for titles
labels <- df %>%
  select(roi, hemisphere, label) %>%
  filter(!is.na(roi)) %>%
  arrange(roi) %>%
  distinct()

# cast factors for plotting consistency
df$effect <- factor(df$effect, c("sentences", "nonwords", "intact speech",
                                 "degraded speech", "hard vWM", "easy vWM"))
df$task <- factor(df$task, c("Lang", "Alice", "MD"))
df$label <- factor(df$label, labels$label %>% unique())

# combined bar plot (left hemisphere only, like your model plots)
df %>%
  mutate(hemilabel = str_c(hemisphere, " ", label)) %>%
  filter(hemisphere == "left", label != "cerebellum") %>%
  ggplot(aes(x = label, y = emmean, ymin = emmean - se, ymax = emmean + se, fill = effect)) +
  geom_col(position = position_dodge(0.9), color = "black") +
  geom_errorbar(position = position_dodge(0.9), linewidth = 1, width = 0) +
  scale_fill_manual(values = cols_fill) +
  scale_y_continuous("Effect size") +
  scale_x_discrete_wrap("parcel") +
  coord_cartesian(ylim = c(-2, 5)) +
  my_theme_text_smaller +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

## save combined plot
ggsave(
  str_c(here("viz", "images", "main", "combined_plots", raw_plot_config$combined_file)),
  dpi = 500, width = 30, height = 16, units = "cm"
)

# make by-ROI plots (each ROI separately)
for (nroi in unique(df$roi)) {

  df %>%
    filter(roi == nroi) %>%
    ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
    geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
    geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
    scale_fill_manual(values = cols_fill) +
    scale_y_continuous("Effect size") +
    scale_x_discrete_wrap("task") +
    guides(fill = FALSE, size = FALSE) +
    ggtitle(str_c(labels$hemisphere[labels$roi == nroi], "\n", labels$label[labels$roi == nroi])) +
    coord_cartesian(ylim = c(-2, 4)) +
    my_theme_text_smaller +
    theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

  ## save each ROI plot
  ggsave(
    here("viz", "images", "main", "by_roi", raw_plot_config$folder,
         str_c("GSS_n772_language_alice_md_raw_roi_", nroi, ".pdf")),
    dpi = 500, width = 5.8, height = 6, units = "cm"
  )
}
```

### Atlas by-roi bar plots (raw data)

```{r}
data_all_atlas_full <- read_csv(here("results", "tables", "atlas_raw_data_detail_all_descriptive_stats.csv")) # ALL participants' raw data, with parcels attached

# helper to load ROI labels and hemisphere info
get_atlas_labels <- function(atlas_type) {
  sheet_name <- switch(atlas_type,
    "DKT" = "DKT",
    "Glasser" = "Glasser",
    "HOCortical" = "HOCortical",
    "HOSubcortical" = "HOSubcortical"
  )
  read_excel(here("data", "AtlasParcels.xlsx"), sheet = sheet_name) %>%
    rename(roi = ROI, ROIName = ROIName, hemi = hemi)
}

# reorder tasks, conditions, and effects
data_all_atlas_full <- data_all_atlas_full %>%
  mutate(
    task = factor(task, levels = c("Lang", "Alice", "MD")),
    condition = factor(condition, levels = c("critical", "baseline")),
    effect = factor(effect, levels = c("sentences", "nonwords", "intact speech", "degraded speech", "hard vWM", "easy vWM"))
    )

# iterate over each atlas
for (atlas_type in unique(data_all_atlas_full$parcels)) {

  df_atlas <- data_all_atlas_full %>% filter(parcels == atlas_type)
  roi_labels <- get_atlas_labels(atlas_type)

  # merge ROI names + hemisphere info
  df_atlas <- df_atlas %>%
    left_join(roi_labels, by = "roi") %>%
    filter(!is.na(ROIName), !is.na(hemi))

  # make atlas-specific folder
  atlas_folder <- switch(atlas_type,
    "DKT" = "atlas_dkt_raw_data_rois",
    "Glasser" = "atlas_glasser_raw_data_rois",
    "HOCortical" = "atlas_hoCort_raw_data_rois",
    "HOSubcortical" = "atlas_hoSubCort_raw_data_rois"
  )
  atlas_plot_dir <- here("viz", "images", "main", "by_roi", atlas_folder)
  dir.create(atlas_plot_dir, recursive = TRUE, showWarnings = FALSE)

  # define hemispheres to loop over
  for (hemisphere in c("left", "right")) {

    df_hemi <- df_atlas %>% filter(str_to_lower(hemi) == hemisphere)
    if (nrow(df_hemi) == 0) next  # skip if empty

    # make hemi-specific subfolder
    hemi_dir <- here(atlas_plot_dir, hemisphere)
    dir.create(hemi_dir, recursive = TRUE, showWarnings = FALSE)

    ## combined plot (faceted by ROI within hemisphere)
    df_hemi %>%
      ggplot(aes(x = task, y = emmean, ymin = emmean - se, ymax = emmean + se, fill = effect)) +
      geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
      geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
      facet_wrap(~ ROIName, scales = "free_x") +
      scale_fill_manual(values = cols_fill) +
      scale_y_continuous("Effect size") +
      coord_cartesian(ylim = c(-2, 5)) +
      ggtitle(str_c("Atlas: ", atlas_type, " | ", str_to_title(hemisphere), " hemisphere")) +
      my_theme_text_smaller +
      theme(
        strip.text = element_text(size = 6),
        axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70)
      )

    ## save combined atlas-hemisphere plot
    ggsave(
      here(atlas_plot_dir, str_c("Atlas_language_alice_md_", atlas_type, "_", hemisphere, "_facet.jpg")),
      dpi = 500, width = 30, height = 20, units = "cm"
    )

    # by-ROI plots (saved to hemi subfolder)
    for (nroi in unique(df_hemi$roi)) {
      this_roi <- df_hemi %>% filter(roi == nroi)
      roi_name <- unique(this_roi$ROIName)

      this_roi %>%
        ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
        geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
        geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
        scale_fill_manual(values = cols_fill) +
        scale_y_continuous("Effect size") +
        scale_x_discrete_wrap("task") +
        guides(fill = FALSE, size = FALSE) +
        ggtitle(str_c(str_to_title(hemisphere), "\n", roi_name)) +
        coord_cartesian(ylim = c(-2, 5)) +
        my_theme_text_smaller +
        theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

      ## save individual ROI plot
      ggsave(
        here(hemi_dir, str_c("Atlas_language_alice_md_", atlas_type, "_", hemisphere, "_", roi_name, "_roi.pdf")),
        dpi = 500, width = 5.8, height = 6, units = "cm"
      )
    }
  }
}
```

### Glasser minimal overlap by-roi bar plots (raw data)

```{r}
# list of parcels of interest with hemisphere
glasser_parcels_of_interest <- tibble(
  ROIName = c("8BL", "OFC", "47m", "TGv", "TE1m", "PeEc", "OP4", "TE2p", "Pir"),
  hemi = c("right", "right", "right", "right", "right", "left", "left", "left", "left")
)

# filter Glasser atlas data
df_glasser <- data_all_atlas_full %>%
  filter(parcels == "Glasser") %>%
  left_join(get_atlas_labels("Glasser"), by = "roi") %>%
  semi_join(glasser_parcels_of_interest, by = c("ROIName", "hemi"))

# print to check the filtered parcels
print(df_glasser %>% select(ROIName, hemi) %>% distinct())

# folder to save plots
glasser_dir <- here("viz", "images", "main", "by_roi", "atlas_glasser_raw_data_rois", "glasser_min_overlap_rois")
dir.create(glasser_dir, recursive = TRUE, showWarnings = FALSE)

# loop over hemisphere
for (hemisphere in c("left", "right")) {
  df_hemi <- df_glasser %>% filter(str_to_lower(hemi) == hemisphere)
  if (nrow(df_hemi) == 0) next
  
  hemi_dir <- here(glasser_dir, hemisphere)
  dir.create(hemi_dir, recursive = TRUE, showWarnings = FALSE)
  
  # combined plot for selected parcels
  df_hemi %>%
    ggplot(aes(x = task, y = emmean, ymin = emmean - se, ymax = emmean + se, fill = effect)) +
    geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
    geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
    facet_wrap(~ ROIName, scales = "free_x") +
    scale_fill_manual(values = cols_fill) +
    coord_cartesian(ylim = c(-1.5, 2.5)) +
    ggtitle(str_c("Glasser minimal-overlap | ", str_to_title(hemisphere), " hemisphere")) +
    my_theme_text_smaller +
    theme(
      strip.text = element_text(size = 6),
      axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70)
    ) -> p_combined
  
  ggsave(
    here(glasser_dir, str_c("Glasser_min_overlap_", hemisphere, "_facet.jpg")),
    plot = p_combined, dpi = 500, width = 30, height = 20, units = "cm"
  )
  
  # individual ROI plots
  for (nroi in unique(df_hemi$roi)) {
    this_roi <- df_hemi %>% filter(roi == nroi)
    roi_name <- unique(this_roi$ROIName)
    
    this_roi %>%
      ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
      geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
      geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
      scale_fill_manual(values = cols_fill) +
      scale_x_discrete_wrap("task") +
      coord_cartesian(ylim = c(-1.5, 2.5)) +
      guides(fill = FALSE, size = FALSE) +
      ggtitle(str_c(str_to_title(hemisphere), "\n", roi_name)) +
      my_theme_text_smaller +
      theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70)) -> p_roi
    
    ggsave(
      here(hemi_dir, str_c("Glasser_min_overlap_", hemisphere, "_", roi_name, "_roi.pdf")),
      plot = p_roi, dpi = 500, width = 5.8, height = 6, units = "cm"
    )
  }
}
```
## Plotting: statistical model data by-ROI bar plots

### GSS cortical / cerebellar bar plots and inset plots (n=772)

```{r}
# define parameters for both sample sizes
plot_configs <- list(
  list(n = 772, df = gss_means_plots_n772, folder = "gss_n772_statistical_models_rois", combined_file = "GSS_n772_statistical_language_alice_md_left_cortical_combined.jpg")
)

for(cfg in plot_configs){

  df <- cfg$df
  
  # ROI names
  labels <- df %>%
    select(roi, hemisphere, label) %>%
    filter(!is.na(roi)) %>%
    arrange(roi) %>%
    unique()
  
  # cast columns as factors for plotting
  df$effect <- factor(df$effect, c("sentences", "nonwords", "intact speech", "degraded speech", "hard vWM", "easy vWM"))
  df$label <- factor(df$label, labels$label %>% unique())
  df$task <- factor(df$task, c("Lang", "Alice", "MD"))
  
  # combined bar plot
  df %>% mutate(hemilabel = str_c(hemisphere, " ", label)) %>% 
    filter(hemisphere=="left", label!="cerebellum") %>% 
    ggplot(aes(x = label, y = emmean, ymin = emmean-se, ymax = emmean+se, fill = effect)) +
    geom_col(position = position_dodge(0.9), color="black") +
    geom_errorbar(position = position_dodge(0.9), linewidth = 1, width = 0) +
    scale_fill_manual(values = cols_fill) +
    scale_y_continuous('Effect size') +
    scale_x_discrete_wrap("parcel") +
    coord_cartesian(ylim = c(-2, 5)) +
    my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))
  
  ## save combined plot
  ggsave(str_c(here("viz", "images", "main", "combined_plots", cfg$combined_file)), dpi=500, width = 30, height = 16, units = "cm")
  
  # by-ROI plots
  for(nroi in 1:27){

    df %>% filter(roi == nroi) %>% 
      ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
      geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
      geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
      scale_fill_manual(values = cols_fill) +
      scale_y_continuous('Effect size') +
      scale_x_discrete_wrap("task") +
      guides(fill = FALSE, size = FALSE) +
      ggtitle(str_c(labels$hemisphere[nroi], "\n", labels$label[nroi])) +
      coord_cartesian(ylim =c(-2, 4)) +
      my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

    ## save plot — here() takes folder and file as separate args
    ggsave(
      here("viz", "images", "main", "by_roi", cfg$folder, str_c("GSS_language_alice_md_", nroi, "_roi_nolegend.pdf")),
      dpi=500, width = 5.8, height = 6, units = "cm"
    )
  }
}

# n=772-specific, language-selective inset plots

# BTLA Cerebellum language-selective: small inset figures
# 12 = BTLA, 24 = right Crus I / II / VIIb, 25 = right VIIb / VIIIa, 23 = left Crus II / VIIb

inset_rois <- c(12, 24, 25, 23)

for(inset_roi in inset_rois){

  gss_means_plots_n772 %>% filter(roi == inset_roi) %>%
    mutate(
      effect = factor(effect, c("sentences", "nonwords", "intact speech", 
                                "degraded speech", "hard vWM", "easy vWM")),
      task   = factor(task, c("Lang", "Alice", "MD"))
    ) %>%
    ggplot(aes(x = task, y=emmean, ymin=lowCI, ymax=highCI, fill=effect)) +
    geom_col(position = position_dodge(0.9), color="black", size=0.25) +
    geom_errorbar(position=position_dodge(0.9), linewidth=0.25, width=0) +
    scale_fill_manual(values=cols_fill) +
    scale_y_continuous(name=NULL, labels=NULL) +
    scale_x_discrete_wrap("task") +
    guides(fill = FALSE, size = FALSE) +
    coord_cartesian(ylim=c(-0.1, 1)) +
    my_theme_text_smaller + theme(axis.text.x = element_text(size = 4, vjust = 1, hjust = 1, angle = 70))

  ## save plot
  ggsave(str_c(here("viz", "images", "main", "by_roi", "gss_n772_statistical_models_rois", "GSS_language_alice_md_"), inset_roi, "_roi_inset.pdf"), 
         dpi=500, width = 2.8, height = 2, units = "cm")
}
```

### Subcortical bar plots (atlas model stats)

```{r}
# subcortical parcels
hosubcortical_parcels <- read_excel(here("data", "AtlasParcels.xlsx"), sheet ="HOSubcortical") %>% rename(roi=ROI)

# subcortical data
subcortical_data <- read.delim(here("results", "stats", "atlas_model_stats", "atlas_roi_models_pairwise_maineffects.txt"), sep = "\t") %>%
  as_tibble() %>%
  # construct error bars with pairwise data, begin by differentiating std. err
  rename("condwise_se" = se)


# provide pairwise std. err
pairwise_se <- read.delim(here("results", "stats", "atlas_model_stats", "atlas_roi_models_pairwise_conditions.txt"), sep = "\t") %>%
  as_tibble() %>%
  filter(atlas == "HOSubcortical") %>%
  select(contrast, se, roi) %>%
  rename("pairwise_se" = se) %>%
  mutate(contrast = case_when(contrast == "Lang critical vs. baseline" ~ "Lang", contrast == "Alice critical vs. baseline" ~ "Alice", contrast == "MD critical vs. baseline" ~ "MD"))

# exclude parcels not of interest (likely show a response to language for reasons other than true sensitivity)
exclude <- c("Left Cerebral Cortex", 
             "Right Cerebral Cortex", 
             "Left Cerebral White Matter", 
             "Right Cerebral White Matter", 
             "Left Lateral Ventricle", 
             "Right Lateral Ventricle",
             "Brain-Stem")

# prepare data
roi_subcortical <- subcortical_data %>% 
  separate(contrast, into=c("task", "condition")) %>%
  mutate(task=ifelse(task=="Language", "Lang", task)) %>% 
  left_join(effects) %>%
  filter(atlas=="HOSubcortical") %>% 
  left_join(hosubcortical_parcels) %>% 
  left_join(lang_responsive_rois) %>% 
  left_join(lang_selective_rois) %>%
  # join pairwise std. err
  left_join(pairwise_se, by = c("roi", "task" = "contrast")) %>% 
  filter(!ROIName%in%exclude) %>% 
  select(1:4, pairwise_se, everything(), -lobe) # reorder columns, drop lobe column

# add contrasts
roi_subcortical$effect <- factor(roi_subcortical$effect, 
                                        c("sentences", 
                                          "nonwords", 
                                          "intact speech", 
                                          "degraded speech", 
                                          "hard vWM", 
                                          "easy vWM"))

# combined bar plot
roi_subcortical %>%
  filter(!is.na(hemi), !ROIName %in% exclude) %>%
  ggplot(aes(x = ROIName, y = estimate, ymin = estimate-pairwise_se, ymax = estimate+pairwise_se, fill = effect)) +
  geom_col(position = position_dodge(0.9), color="black") +
  geom_errorbar(position = position_dodge(0.9), linewidth = 1, width = 0) +
  scale_fill_manual(values = cols_fill) +
  scale_y_continuous('Effect size') +
  scale_x_discrete_wrap("task") +
  facet_grid(~hemi, scales = "free_x") +
  my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

## save plot
ggsave(str_c(here("viz", "images", "main", "combined_plots", "HOSubcortical_statistical_language_alice_md_left_right_cortical_combined.jpg")), dpi=500, width = 30, height = 16, units = "cm")

# 7 - write data
write.table(roi_subcortical, here("results", "tables", "HOSubcortical_results.txt"), quote=FALSE, row.names = FALSE, sep = "\t")

# subcortical labels
subcort_rois <- roi_subcortical %>% select(roi) %>% unique() %>% pull()
subcort_labels <- roi_subcortical %>% select(roi, ROIName) %>% filter(roi%in%subcort_rois) %>% arrange(roi) %>% unique()

# by-ROI subcortical plots
n=0
for(nroi in subcort_rois){ # uses within groups
  n=n+1
  roi_subcortical %>%
    filter(roi == nroi) %>%
    mutate(task = factor(task, levels = c("Lang", "Alice", "MD"))) %>%  # set order
    ggplot(aes(x = task, y = estimate, ymin = estimate-pairwise_se, ymax = estimate+pairwise_se, fill = effect)) +
    geom_col(position = position_dodge(0.9), color = "black", size=0.25) +
    geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
    scale_fill_manual(values = cols_fill) +
    scale_y_continuous('Effect size') +
    scale_x_discrete_wrap("task") +
    guides(fill = FALSE, size = FALSE) +
    ggtitle(subcort_labels$ROIName[n]) +
    coord_cartesian(ylim = c(-0.2, 0.55)) +
    my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))
  
  ## save plot
  ggsave(str_c(here("viz", "images", "main", "by_roi", "atlas_hoSubCort_statistical_models_rois", "HOSub_language_alice_md_"), nroi, "_roi_nolegend.pdf"), dpi=500, width = 5.8, height = 6, units = "cm")
}
```

## Supplemental plotting

### Parcel overlap histograms

See supplementary information.

```{r}
# quantifications of the overlap percentage between the atlas and functional language parcels

# DKT
atlas_overlap_average %>% filter(parcels == "DKT") %>%
  ggplot(aes(x = OverlapPercentage)) + 
  geom_histogram(bins = 10, aes(y = (..count..) / sum(..count..) * 100), fill = "indianred3", color = "black") +
  scale_y_continuous("Percent of parcels \nshowing a given overlap") +
  scale_x_continuous("Percent overlap\nwith functional parcels") +
  coord_cartesian(ylim = c(0,60), xlim = c(-5, 105))+
  my_theme_text_smaller
ggsave(here("viz", "images", "main", "by_atlas", "dkt", "DKT_overlap_histogram.pdf"), dpi=500, width = 8, height = 6, units = "cm")

# Glasser
atlas_overlap_average %>% filter(parcels == "Glasser") %>%
  ggplot(aes(x = OverlapPercentage)) + 
  geom_histogram(bins = 10, aes(y = (..count..) / sum(..count..) * 100), fill = "indianred3", color = "black") +
  scale_y_continuous("Percent of parcels \nshowing a given overlap") +
  scale_x_continuous("Percent overlap\nwith functional parcels") +
  coord_cartesian(ylim = c(0,60), xlim = c(-5, 105))+
  my_theme_text_smaller
ggsave(here("viz", "images", "main", "by_atlas", "glasser", "Glasser_overlap_histogram.pdf"), dpi=500, width = 8, height = 6, units = "cm")

# HOC
atlas_overlap_average %>% filter(parcels == "HOC") %>%
  ggplot(aes(x = OverlapPercentage)) + 
  geom_histogram(bins = 10, aes(y = (..count..) / sum(..count..) * 100), fill = "indianred3", color = "black") +
  scale_y_continuous("Percent of parcels \nshowing a given overlap") +
  scale_x_continuous("Percent overlap\nwith functional parcels") +
  coord_cartesian(ylim = c(0,60), xlim = c(-5, 105))+
  my_theme_text_smaller
ggsave(here("viz", "images", "main", "by_atlas", "hoCort", "HOC_overlap_histogram.pdf"), dpi=500, width = 8, height = 6, units = "cm")
```

### Atlas plots for figures: Binary, responsive, and selective brain plots (with and without annotations)

The braincollage() function is one of two functions sourced from the plot_brains.R script. Braincollage() is a wrapper for brainplot(), a complex function with pre-defined aesthetics and renderings. If intended to change the visual output, make changes to brainplot() and resource.

```{r}
#| include: false

# create a list of images for each plot (from source plot_brains.R)

# DKT
braincollage(
  df = dkt_plot_lang,  # df for aseg data
  atlas = dkt  # ggseg atlas object for aseg
)

# Glasser
braincollage(
  df = glasser_plot_lang,  # df for aseg data
  atlas = glasser  # ggseg atlas object for aseg
)

# HOC
braincollage(
  df = hoCort_plot_lang,  # df for aseg data
  atlas = hoCort  # ggseg atlas object for aseg
)

# HOSC
braincollage(
  df = hoSubCort_plot_lang,  # df for aseg data
  atlas = aseg  # ggseg atlas object for aseg
)
```

### Atlas plots for figures: Main cortical figure (base)

Populates the base images comprising the text's main cortical figure. This does not create the figure itself.

```{r}
#| eval: false
### Layer 1

p1 <- brainplot(
      df = dkt_plot_lang,
      atlas = dkt,
      plot_type = "binary_dual",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "all",
      folder = "main_cortical_figure",
      file_name = "dkt"
    )

p2 <- brainplot(
      df = hoCort_plot_lang,
      atlas = hoCort,
      plot_type = "binary_dual",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "all",
      folder = "main_cortical_figure",
      file_name = "hoCort"
    ) 

p3 <- brainplot(
      df = glasser_plot_lang,
      atlas = glasser,
      plot_type = "binary_dual",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "all",
      folder = "main_cortical_figure",
      file_name = "glasser"
    )


### Layer 2 - Lang
p4 <- brainplot(
      df = dkt_plot_lang,
      atlas = dkt,
      plot_type = "selective",
      task = "Lang",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "dkt_lang"
    )

p5 <- brainplot(
      df = hoCort_plot_lang,
      atlas = hoCort,
      plot_type = "selective",
      task = "Lang",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "hocort_lang"
    )

p6 <- brainplot(
      df = glasser_plot_lang,
      atlas = glasser,
      plot_type = "selective",
      task = "Lang",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "glasser_lang"
    )


### Layer 3 - Alice
p7 <- brainplot(
      df = dkt_plot_lang,
      atlas = dkt,
      plot_type = "selective",
      task = "Alice",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "dkt_alice"
    )

p8 <- brainplot(
      df = hoCort_plot_lang,
      atlas = hoCort,
      plot_type = "selective",
      task = "Alice",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "hocort_alice"
    )

p9 <- brainplot(
      df = glasser_plot_lang,
      atlas = glasser,
      plot_type = "selective",
      task = "Alice",
      plot_title = "",
      subtitle = NULL,
      legend_title = FALSE,
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "glasser_alice"
    )


### Layer 4 - MD
p10 <- brainplot(
      df = dkt_plot_lang,
      atlas = dkt,
      plot_type = "selective",
      task = "MD",
      plot_title = "",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "dkt_md"
    )

p11 <- brainplot(
      df = hoCort_plot_lang,
      atlas = hoCort,
      plot_type = "selective",
      task = "MD",
      plot_title = "",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "hocort_md"
    )

p12 <- brainplot(
      df = glasser_plot_lang,
      atlas = glasser,
      plot_type = "selective",
      task = "MD",
      plot_title = "",
      subtitle = NULL,
      legend_title = "Activation",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      view = "left_lateral",
      folder = "main_cortical_figure",
      file_name = "glasser_md"
    )

cortical_patch_plot <-  (p1 | p2 | p3) / (p4 | p5 | p6) / (p7 | p8 | p9) / (p10 | p11 | p12)
  
    ggsave(str_c(here("viz", "images", "main", "by_atlas", "main_cortical_figure", "main_cortical_figure_base"), ".pdf"), plot = cortical_patch_plot, dpi=500, width = 42, height = 48, units = "in")  # default: 18 x 14; patchwork: 32 x 26
```

### Atlas plots for figures: Main subcortical figure (base)

Likewise, this only populates images included in the main subcortical figure within the text, *not* the full figure itself.

```{r}
#| eval: false
subp1 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Amygdala",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      outline_regions = c("amygdala"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg1"
    )

subp2 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Thalamus",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      outline_regions = c("thalamus proper"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg2"
    )

subp3 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Hippocampus",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = TRUE,
      outline_regions = c("hippocampus"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg3"
    )

subp4 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Putamen",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = FALSE,
      # outline_regions = c("putamen"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg4"
    )

subp5 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Pallidum",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = FALSE,
      # outline_regions = c("pallidum"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg5"
    )

subp6 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Caudate",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = FALSE,
      # outline_regions = c("caudate"),
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg6"
    )

subp7 <- brainplot(
      df = hoSubCort_plot_lang,
      atlas = aseg,
      plot_type = "binary_dual",
      plot_title = "Accumbens",
      subtitle = NULL,
      legend_title = "",
      legend_on = FALSE,
      xlab = "",
      ylab = "",
      annotate = FALSE,
      outline = FALSE,
      view = "coronal",
      folder = "main_subcortical_figure",
      file_name = "aseg7"
    )

subcortical_patch_plot <- (plot_spacer() | subp1 | plot_spacer()) / 
  (plot_spacer() | subp2 | plot_spacer()) /
  (plot_spacer() | subp3 | plot_spacer()) /
  (plot_spacer() | subp4 | plot_spacer()) /
  (plot_spacer() | subp5 | plot_spacer()) /
  (plot_spacer() | subp6 | plot_spacer()) /
  (plot_spacer() | subp7 | plot_spacer())
  
    ggsave(str_c(here("viz", "images", "main", "by_atlas", "main_subcortical_figure", "main_subcortical_figure_base"), ".pdf"), plot = subcortical_patch_plot, dpi=500, width = 42, height = 48, units = "in")  # default: 18 x 14; patchwork: 32 x 26
```