---
title: "Extended Language Network | 13_n772_n86_model_comparisons.qmd"
author: "Aaron Wright [`aawva@mit.edu`], Agata Wolna [`awolna@mit.edu`]"
date: "2025-10-23"
format: html
---

# Overview

**Purpose**: run statistical models for gss and atlas parcels, and create counts of language-selective and language-responsive rois

**Input(s)**: `detail_all.txt`, the main data table

**Process**: run models for the gss n=772, gss n=86, and atlas parcels; describe the distribution of parcels; calculate roi sizes, and create tables of lang-selective and lang-responsive parcels

**Output(s)**: restults/stats/`atlas_model_stats` \| `gss_n772_model_stats` \| `gss_n86_model_stats`: lists of model outputs and optional diagnostics; `atlas_lang_responsive_rois_names.txt` and `atlas_lang_selective_rois_names.txt`: descriptive tables for atlas parcels

# Environment

```{r}
library(tidyverse)
library(here)
library(readxl)
library(lmerTest)
library(emmeans)

source(here("scripts", "run_parcel_models.R"))
source(here("scripts", "plotting_aesthetics.R")) # EvLab contrast aesthetics
```

This pipeline is organized into two sections, stats and plotting, to be consistent with the pipeline adopted for the GSS n772 and atlas aprcel models in 02_stats.qmd and 03_plots.qmd. This script is a acollapsing of the goals of those two scripts as applied to the GSS n86 model.

# Section I : Stats pipeline

## Data (raw data)

```{r}
# read in the data
# read in the data
data_all <- read.delim(here("data", "detail_all.txt")) # ALL participants

# create a data frame to define task effects
effects <- data.frame(
  column1 = c("Alice", "Lang", "MD", "Alice", "Lang", "MD"),
  column2 = c("baseline", "baseline", "baseline", "critical", "critical", "critical"),
  column3 = c("degraded speech", "nonwords", "easy vWM", "intact speech", "sentences", "hard vWM"))
# rename columns
colnames(effects) <- c("task", "condition", "effect")

# read in parcel decoding files
parcels <- read.delim(here("data", "recode_parcels.txt"), sep = "\t") %>% select(-ROI) %>% rename("roi" = "ROI_new")
```

## Parcel stats

Group-constrained specific-subject parcellation method (see Fedorenko et al., 2010 for a description of the approach).

## Tidying and raw data descriptive statistics

Following the statistical model approach outlined in 02_stats for the GSS n772 model, we compute statistics for a subset of the parcels as defined by individuals who completed all three localizer tasks.

```{r}
#| eval: false

gcssroi_excluded <- c(9, 29, 30, 31, 32, 33) # exclude unnecessary parcels before recoding

# recode parcels with new ROI numbers
data_gcss <- data_all %>% 
  mutate(condition = as.factor(condition), task = as.factor(task)) %>%
  filter(parcels =="GcSS", !(ROI %in% gcssroi_excluded)) # filter for only desired ROIs

gss_parcels_recoded <- read.delim(here("data", "recode_parcels.txt"), sep = "\t") %>% # recode parcels with new ROI numbers
  filter(!(ROI %in% gcssroi_excluded)) %>%
  mutate(
    atlas = "GSS",
    lobe = NA_character_) %>%
  rename(
    hemi = hemisphere,
    ROIName = label)

data_gcss_recoded <- data_gcss %>% left_join(gss_parcels_recoded, by = "ROI") %>%
  select(-ROI) %>% # drop old parcel numbers
  rename(ROI = ROI_new) # rename new, recoded parcel numbers

# subset for data of participants having completed all three tasks (does not yet compute new df)
data_gcss_uid <- data_gcss_recoded %>% filter(localizer!="Alice") %>% 
  separate(Subject, into = "uid", sep = "_", remove = FALSE)

# data, secondary n check 
data_gcss_uid %>%
  select(uid, task, localizer) %>%
  unique() %>%
  group_by(uid) %>%
  count() %>% 
  group_by(n) %>%
  count()

# subset list of participants having completed all three tasks (applies filtering)
uids_n86 <- data_gcss_uid %>%
  select(uid, task, localizer) %>%
  unique() %>%
  mutate(compl=1) %>% 
  pivot_wider(names_from = task, values_from = compl) %>%
  filter(Lang==1 & MD==1 & Alice == 1) %>%
  pull(uid)

# set df to be used in n=86 parcel model
data_gcss_n86 <- data_gcss_uid %>%
  filter(uid %in% uids_n86)

# data, validation n check
data_gcss_n86 %>%
  select(uid, task, localizer) %>%
  unique() %>%
  group_by(uid) %>%
  count() %>% 
  group_by(n) %>%
  count()

write_csv(data_gcss_n86, here("results", "tables", "gss_raw_data_detail_n86_recoded.csv")) # ALL participants' raw data, with parcels attached

# join with effects + parcels
data_n86_recoded_full <- data_gcss_n86 %>%
  left_join(effects, by = c("task", "condition"))
  # rename(roi = ROI, hemisphere = hemi, label = ROIName)

data_n86_recoded_gss_full <- data_n86_recoded_full %>%
  rename(
    roi = ROI,
    hemisphere = hemi,
    label = ROIName
    ) %>%
  group_by(task, condition, roi, hemisphere, label, effect) %>%
  summarise(
    emmean = mean(EffectSize, na.rm = TRUE),
    se     = sd(EffectSize, na.rm = TRUE) / sqrt(n()),
    lowCI  = emmean - qt(0.975, df = n() - 1) * se,
    highCI = emmean + qt(0.975, df = n() - 1) * se,
    .groups = "drop"
  ) %>%
  mutate(
    atlas = "GSS",
    model = "n=86",
    # enforce factor order to control sorting
    task = factor(task, levels = c("Alice", "Lang", "MD")),
    condition = factor(condition, levels = c("baseline", "critical"))
  ) %>%
  # order the rows within each roi in your desired sequence
  arrange(roi, hemisphere, label, condition, task) %>%
  relocate(task, condition, emmean, se, lowCI, highCI, roi, atlas, effect, hemisphere, label, model)

write_csv(data_n86_recoded_gss_full, here("results", "tables", "gss_raw_data_detail_n86_recoded_descriptive_stats.csv")) # ALL participants' raw data, with parcels attached

```

## Parcel model: gss parcels n=86 (subset)

```{r}
# run parcel model (gss, n=86)
run_parcel_models(data_gcss_n86,
                  parcels_filter="GcSS",
                  target_data="gss_n86",
                  diagnostics=FALSE)
```

## Parcel tables: statistical model output

### Aggregate data for computing tables (statistical model data)

```{r}
# read in the data

# gss n86
gss86_maineffects <- read.table(here("results", "stats", "gss_n86_model_stats", "gss_n86_roi_models_pairwise_maineffects.txt"), sep = "\t", header = TRUE) %>% as_tibble()
gss86_conditions <- read.table(here("results", "stats", "gss_n86_model_stats", "gss_n86_roi_models_pairwise_conditions.txt"), sep = "\t", header = TRUE) %>% as_tibble()
gss86_tasks <- read.table(here("results", "stats", "gss_n86_model_stats", "gss_n86_roi_models_pairwise_tasks.txt"), sep = "\t", header = TRUE) %>% as_tibble()
```

### Language-selective and language-responsive ROIs (GSS n=86)

```{r}
gcssroi_excluded <- c(9, 29, 30, 31, 32, 33) # exclude unnecessary parcels before recoding

gss_parcels_recoded <- read.delim(here("data", "recode_parcels.txt"), sep = "\t") %>% # recode parcels with new ROI numbers
  filter(!(ROI %in% gcssroi_excluded)) %>%
  mutate(
    atlas = "GSS",
    lobe = NA_character_) %>%
  rename(
    hemi = hemisphere,
    ROIName = label)

# Repeat for n=86
gss86_maineffects_rois <- gss86_maineffects %>%
  separate(contrast, into=c("task", "effect")) %>%
  filter(effect=="critical", task!="MD", estimate>0, p_value<0.05) %>%
  group_by(roi) %>%
  count() %>%
  filter(n==2) %>%
  mutate(maineffect=1) %>%
  select(-n)

gss86_responsive_rois <- gss86_conditions %>%
  separate(contrast, into=c("task")) %>%
  filter(task!="MD", p_value<0.05) %>%
  group_by(roi) %>%
  count() %>%
  filter(n==2) %>%
  select(-n) %>%
  left_join(gss86_maineffects_rois, by="roi") %>%
  filter(!is.na(maineffect)) %>%
  select(-maineffect) %>%
  unique()

gss86_selective_rois <- gss86_tasks %>%
  separate(contrast, into=c("task1", "task2"), sep = " vs. ") %>%
  filter(task2 == "criticalMD", estimate>0, p_value<0.05) %>%
  group_by(roi) %>%
  count() %>%
  filter(n==2) %>%
  select(-n) %>%
  mutate(selective=1)

gss86_selective_rois <- gss86_responsive_rois %>%
  left_join(gss86_selective_rois, by="roi") %>%
  filter(!is.na(selective)) %>%
  select(-selective)

gss86_responsive_rois <- gss86_responsive_rois %>%
  left_join(gss_parcels_recoded %>% select(ROI=ROI_new, ROIName, hemi), by=c("roi"="ROI")) %>%
  unique()
gss86_selective_rois <- gss86_selective_rois %>%
  left_join(gss_parcels_recoded %>% select(ROI=ROI_new, ROIName, hemi), by=c("roi"="ROI")) %>%
  unique()

write.table(gss86_responsive_rois, here("results", "tables", "gss86_lang_responsive_rois_names.txt"), sep = "\t", row.names = FALSE, quote = FALSE)
write.table(gss86_selective_rois, here("results", "tables", "gss86_lang_selective_rois_names.txt"), sep = "\t", row.names = FALSE, quote = FALSE)
```

### (Optional) Check: Language-selective and language-responsive ROIs

```{r}
gss86_responsive_rois %>% 
  mutate(atlas = "GSS_n86") %>% 
  group_by(atlas, hemi) %>% 
  count()

gss86_selective_rois %>% 
mutate(atlas = "GSS_n86") %>% 
  group_by(atlas, hemi) %>% 
  count()
```

# Section II : Plotting pipeline

## Data (raw data)

```{r}
data_gcss_n86 <- read_csv(here("results", "tables", "gss_raw_data_detail_n86_recoded_descriptive_stats.csv")) # ALL participants' raw data, with parcels attached (n=86)
```

## Plotting: raw data by-ROI bar plots (n=86)

### GSS cortical by-roi bar plots (raw data) (n=86)

```{r}
# read in the data

# define config for raw pipeline
raw_plot_config <- list(
  df = data_gcss_n86,
  folder = "gss_n86_raw_data_rois",
  combined_file = "GSS_n86_raw_language_alice_md_left_cortical_combined.jpg"
)

df <- raw_plot_config$df

# roi labels for titles
labels <- df %>%
  select(roi, hemisphere, label) %>%
  filter(!is.na(roi)) %>%
  arrange(roi) %>%
  distinct()

# cast factors for plotting consistency
df$effect <- factor(df$effect, c("sentences", "nonwords", "intact speech",
                                 "degraded speech", "hard vWM", "easy vWM"))
df$task <- factor(df$task, c("Lang", "Alice", "MD"))
df$label <- factor(df$label, labels$label %>% unique())

# combined bar plot (left hemisphere only, like your model plots)
df %>%
  mutate(hemilabel = str_c(hemisphere, " ", label)) %>%
  filter(hemisphere == "left", label != "cerebellum") %>%
  ggplot(aes(x = label, y = emmean, ymin = emmean - se, ymax = emmean + se, fill = effect)) +
  geom_col(position = position_dodge(0.9), color = "black") +
  geom_errorbar(position = position_dodge(0.9), linewidth = 1, width = 0) +
  scale_fill_manual(values = cols_fill) +
  scale_y_continuous("Effect size") +
  scale_x_discrete_wrap("parcel") +
  coord_cartesian(ylim = c(-2, 5)) +
  my_theme_text_smaller +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

## save combined plot
ggsave(
  str_c(here("viz", "images", "supplemental", "combined_plots", raw_plot_config$combined_file)),
  dpi = 500, width = 30, height = 16, units = "cm"
)

# make by-ROI plots (each ROI separately)
for (nroi in unique(df$roi)) {

  df %>%
    filter(roi == nroi) %>%
    ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
    geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
    geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
    scale_fill_manual(values = cols_fill) +
    scale_y_continuous("Effect size") +
    scale_x_discrete_wrap("task") +
    guides(fill = FALSE, size = FALSE) +
    ggtitle(str_c(labels$hemisphere[labels$roi == nroi], "\n", labels$label[labels$roi == nroi])) +
    coord_cartesian(ylim = c(-2, 4)) +
    my_theme_text_smaller +
    theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

  ## save each ROI plot
  ggsave(
    here("viz", "images", "supplemental", "by_roi", raw_plot_config$folder,
         str_c("GSS_language_alice_md_raw_", nroi, "_roi.pdf")),
    dpi = 500, width = 5.8, height = 6, units = "cm"
  )
}
```

## Statisical models

```{r}
# read in data

# GSS model - full model (for model comparison plots)
gss_means_plots_n772 <- read.delim(here("results", "stats", "gss_n772_model_stats", "gss_n772_roi_models_all_means.txt")) %>% as_tibble() %>% select(-df)

# GSS model - n86 (only participants with all 3 tasks)
gss_means_plots_n86 <- read.delim(here("results", "stats", "gss_n86_model_stats", "gss_n86_roi_models_all_means.txt")) %>% as_tibble() %>% select(-df)

# join effects + gss_means_plots - n772
gss_means_plots_n772 <- gss_means_plots_n772 %>%
  left_join(effects, by = c("task", "condition")) %>%
  left_join(parcels, by = "roi") %>%
  mutate(model = "n=772")

# join effects + gss_means_plots - n86
gss_means_plots_n86 <- gss_means_plots_n86 %>%
  left_join(effects, by = c("task", "condition")) %>%
  left_join(parcels, by = "roi") %>%
  mutate(model = "n=86", atlas = "GSS")
```

## Plotting: statistical model data by-ROI bar plots

### GSS cortical / cerebellar bar plots (n=86)

```{r}
plot_configs <- list(
  list(n = 86, df = gss_means_plots_n86, folder = "gss_n86_statistical_models_rois", combined_file = "GSS_n86_statistical_language_alice_md_left_cortical_combined.jpg")
)

for(cfg in plot_configs){

  df <- cfg$df
  
  # ROI names
  labels <- df %>%
    select(roi, hemisphere, label) %>%
    filter(!is.na(roi)) %>%
    arrange(roi) %>%
    unique()
  
  # cast columns as factors for plotting
  df$effect <- factor(df$effect, c("sentences", "nonwords", "intact speech", "degraded speech", "hard vWM", "easy vWM"))
  df$label <- factor(df$label, labels$label %>% unique())
  df$task <- factor(df$task, c("Lang", "Alice", "MD"))
  
  # combined bar plot
  df %>% mutate(hemilabel = str_c(hemisphere, " ", label)) %>% 
    filter(hemisphere=="left", label!="cerebellum") %>% 
    ggplot(aes(x = label, y = emmean, ymin = emmean-se, ymax = emmean+se, fill = effect)) +
    geom_col(position = position_dodge(0.9), color="black") +
    geom_errorbar(position = position_dodge(0.9), linewidth = 1, width = 0) +
    scale_fill_manual(values = cols_fill) +
    scale_y_continuous('Effect size') +
    scale_x_discrete_wrap("parcel") +
    coord_cartesian(ylim = c(-2, 5)) +
    my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))
  
  ## save combined plot
  ggsave(str_c(here("viz", "images", "supplemental", "combined_plots", cfg$combined_file)), dpi=500, width = 30, height = 16, units = "cm")
  
  # by-ROI plots
  for(nroi in 1:27){

    df %>% filter(roi == nroi) %>% 
      ggplot(aes(x = task, y = emmean, ymin = lowCI, ymax = highCI, fill = effect)) +
      geom_col(position = position_dodge(0.9), color = "black", size = 0.25) +
      geom_errorbar(position = position_dodge(0.9), linewidth = 0.25, width = 0) +
      scale_fill_manual(values = cols_fill) +
      scale_y_continuous('Effect size') +
      scale_x_discrete_wrap("task") +
      guides(fill = FALSE, size = FALSE) +
      ggtitle(str_c(labels$hemisphere[nroi], "\n", labels$label[nroi])) +
      coord_cartesian(ylim =c(-2, 4)) +
      my_theme_text_smaller + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))

    ## save plot — here() takes folder and file as separate args
    ggsave(
      here("viz", "images", "supplemental", "by_roi", cfg$folder, str_c("GSS_language_alice_md_", nroi, "_roi_nolegend.pdf")),
      dpi=500, width = 5.8, height = 6, units = "cm"
    )
  }
}
```

### Full model n=772 vs. partial model n=86 model plots

```{r}
# set comparison df
means_all_n86_comparison <- rbind(gss_means_plots_n772, gss_means_plots_n86)

# set ROI labels
labels <- means_all_n86_comparison %>% select(roi, hemisphere, label) %>% filter(!is.na(roi)) %>% arrange(roi) %>% unique()

means_all_n86_comparison$effect <- factor(means_all_n86_comparison$effect, c("sentences", "nonwords", "intact speech", "degraded speech", "hard vWM", "easy vWM"))
means_all_n86_comparison$label <- factor(means_all_n86_comparison$label, labels$label %>% unique())
means_all_n86_comparison$task <- factor(means_all_n86_comparison$task, c("Lang", "Alice", "MD"))

# by-ROI GSS
for(nroi in 1:27){
  means_all_n86_comparison %>% filter(roi == nroi) %>%
    ggplot(aes(x = task, y=emmean, ymin=lowCI, ymax=highCI, fill=effect)) +
    geom_col(position = position_dodge(0.9), color="black", size=0.25) +
    geom_errorbar(position=position_dodge(0.9), linewidth=0.25, width=0) +
    scale_fill_manual(values=cols_fill) +
    scale_y_continuous('Effect size') +
    scale_x_discrete_wrap("task") +
    guides(fill = FALSE, size = FALSE) +
    facet_grid(~model, scales="free_x") +
    ggtitle(str_c(labels$hemisphere[nroi], "\n", labels$label[nroi])) +
    coord_cartesian(ylim=c(-2, 5)) +
    my_theme_text_smaller_noborders + theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 70))
  ggsave(str_c(here("viz", "images", "supplemental", "by_roi", "gss_n772_n86_statistical_models_comparison", "GCSS_language_alice_md_"), nroi, "_roi_nolegend.pdf"), dpi=500, width = 5.8, height = 6, units = "cm")
}
```